IMF and Benchmark Forecasts

### 2023-10-17 manuscript/manuscript.Rmd
move more functions to designated plot file

ovr_assessment_plot <- function(wis_scoredat, cvg_scoredat, point_scores, crps_scores, trgt){

  cvgscores <- cvg_scoredat |>
    .d(method == "rolling window") |>
    .d(target == trgt) |>
    .d(, target := NULL) |>
    .d(, method := NULL) #|>

  decomp_data <- wis_scoredat |>
    .d(method == "rolling window") |>
    .d(target == trgt) |>
    .d(, target := NULL) |>
    .d(, method := NULL) #|>

@@

  crps_scores_rw <- crps_scores |>
    .d(method == "rolling window") |>
    .d(target == trgt) |>
    .d(, target := NULL)

@@

  xmax <- max(decomp_data$interval_score)

  decomp_data <- decomp_data |>
    data.table::melt(id.vars = c("source", "error_method", "horizon"),
                     value.name = "score",
                     variable.name = "type") |>
    .d(type != "interval_score") # filter out total score, only interested in decomposition

  #check arguments of function!
  wis_plot <- function(decompdat, error_m, maxval){

    if(error_m == "directional"){
      plot_name <- "WIS - Directional Errors"
    } else {
      plot_name <- "WIS - Absolute Errors"
    }

    plt <- ggplot(decomp_data |> .d(error_method == error_m),
                          aes(x = source, 
                              fill = source,
                              y = score,
                              alpha = type)) +
      geom_bar(position="stack", stat="identity") +
      facet_wrap(~ horizon,
                 labeller = as_labeller(plot_horizon_label_short(4)),
                 nrow = 1) +
      #geom_hline(yintercept = 1) +
      scale_alpha_discrete(range = c(0.35,1),
                           labels = c("Overprediction",
                                      "Underprediction",
                                      "Dispersion"),
                           name = "Components of the WIS") +
      scale_x_discrete(name = "",
                       breaks = c("IMF", "bvar", "bvar_qu", "ar"),
                       labels = c("IMF", "BVAR", "BVAR_QU", "AR")) +
      scale_y_continuous(name = "WIS", limits = c(0, maxval)) +
      scale_fill_met_d("Hokusai1") +
      #scale_fill_brewer(palette = "Set2", 
      #                  name = "Model Type")+
      xlab("") +
      #labs(fill = "Model Type") +
      theme_uqimf() %+replace%
      theme(axis.text.x = element_text(size = 8, angle = 90, hjust = .5, vjust = .5, face = "plain"),
            strip.text = element_text(size = 8)) +
      ggtitle(plot_name) +
      theme(#legend.title=element_blank(),
            legend.position = "none")

    return(plt)

  }

  horizon_path_plot <- function(score_data,
                                score_rule,
                                ylab = score_rule, 
                                facet = NULL,
                                title = NULL){

    path_plot <- ggplot(score_data, 
           aes(x = horizon,
               y = get(score_rule), 
               color = source)) +

      scale_color_met_d("Hokusai1", breaks = c("IMF", "bvar", "ar", "bvar_qu"),
                       labels = c("IMF", "BVAR", "AR", "BVAR_QU")) +

      #scale_color_brewer(palette = "Set2") +
      geom_line() +
      geom_point() +
      ylab(ylab) +
      xlab("Forecast Horizon") +
      guides(fill = "none", colour = "none") +
      theme_uqimf() %+replace%
      theme(legend.title=element_blank(),
            legend.position = "none") +
      ggtitle(title)

    if(!is.null(facet)){

      path_plot <- path_plot +
      facet_wrap(~ get(facet),
                 scales = "free") 
    }

    return(path_plot)

  }

  dir_wis_plot <- wis_plot(decomp_data, "directional", xmax)
  abs_wis_plot <- wis_plot(decomp_data, "absolute", xmax)

@@

  crpsplot_rw <- horizon_path_plot(crps_scores_rw |>.d(, error_method := "directional"), "score", "CRPS by Sample", title = "CRPS - Rolling Window") +
    scale_y_continuous(limits = c(0.07,1.05))

@@

ovr_assessment_plot(wis_scores, scores_cvgshort, point_scores, all_scores, "pcpi_pch")

```
\newpage
Some notes:
\begin{itemize}
\item Inflation: directional vs. absolute errors:
\begin{itemize}
\item difference small for IMF method, absolute slighly better, likely due to longer central intervals
\item AR and BVAR profit more from directional correction (upward bias)
\item for expanding window method, difference in coverage is smaller (-> structural breaks)

@@

```{r gdpovr_ew, echo = FALSE, warning = TRUE, fig.dim=c(12,12)}

ovr_assessment_plot(wis_scores, scores_cvgshort, point_scores, all_scores, "ngdp_rpch")
```

@@

```{r overall_assessment, fig.dim=c(12,12), echo = FALSE, warning = FALSE}

bvar_scores_cvgshort <- fread(here("scores", "bvar_ci_scores_avgcnt.csv")) |>
  setnames("model", "source", skip_absent = TRUE) |>
  .d(, source := "bvar_qu") |>
  .d(, method := "expanding window") |>
  .d(, error_method := "absolute")

ovr_assessment_plot <- function(wis_scoredat, cvg_scoredat, point_scores, crps_scores, trgt){

  cvgscores <- cvg_scoredat |>
    .d(method == "expanding window") |>
    .d(target == trgt) |>
    .d(, target := NULL) |>
    .d(, method := NULL) #|>

  decomp_data <- wis_scoredat |>
    .d(method == "expanding window") |>
    .d(target == trgt) |>
    .d(, target := NULL) |>
    .d(, method := NULL) #|>

  crps_scores_rw <- crps_scores |>
    .d(method == "expanding window") |>
    .d(target == trgt) |>
    .d(, target := NULL)

  xmax <- max(decomp_data$interval_score)

  decomp_data <- decomp_data |>
    data.table::melt(id.vars = c("source", "error_method", "horizon"),
                     value.name = "score",
                     variable.name = "type") |>
    .d(type != "interval_score") # filter out total score, only interested in decomposition

  #check arguments of function!
  wis_plot <- function(decompdat, error_m, maxval){

    if(error_m == "directional"){
      plot_name <- "WIS - Directional Errors"
    } else {
      plot_name <- "WIS - Absolute Errors"
    }

    plt <- ggplot(decomp_data |> .d(error_method == error_m),
                          aes(x = source, 
                              fill = source,
                              y = score,
                              alpha = type)) +
      geom_bar(position="stack", stat="identity") +
      facet_wrap(~ horizon,
                 labeller = as_labeller(plot_horizon_label_short(4)),
                 nrow = 1) +
      #geom_hline(yintercept = 1) +
      scale_alpha_discrete(range = c(0.35,1),
                           labels = c("Overprediction",
                                      "Underprediction",
                                      "Dispersion"),
                           name = "Components of the WIS") +
      scale_x_discrete(name = "",
                       breaks = c("IMF", "bvar", "bvar_qu", "ar"),
                       labels = c("IMF", "BVAR", "BVAR_QU", "AR")) +
      scale_y_continuous(name = "WIS", limits = c(0, maxval)) +
      scale_fill_met_d("Hokusai1") +
      #scale_fill_brewer(palette = "Set2", 
      #                  name = "Model Type")+
      xlab("") +
      #labs(fill = "Model Type") +
      theme_uqimf() %+replace%
      theme(axis.text.x = element_text(size = 8, angle = 90, hjust = .5, vjust = .5, face = "plain"),
            strip.text = element_text(size = 8)) +
      ggtitle(plot_name) +
      theme(#legend.title=element_blank(),
            legend.position = "none")

    return(plt)

  }

  horizon_path_plot <- function(score_data,
                                score_rule,
                                ylab = score_rule, 
                                facet = NULL,
                                title = NULL){

    path_plot <- ggplot(score_data, 
           aes(x = horizon,
               y = get(score_rule), 
               color = source)) +

      scale_color_met_d("Hokusai1", breaks = c("IMF", "bvar", "ar", "bvar_qu"),
                       labels = c("IMF", "BVAR", "AR", "BVAR_QU")) +

      #scale_color_brewer(palette = "Set2") +
      geom_line() +
      geom_point() +
      ylab(ylab) +
      xlab("Forecast Horizon") +
      guides(fill = "none", colour = "none") +
      theme_uqimf() %+replace%
      theme(legend.title=element_blank(),
            legend.position = "none") +
      ggtitle(title)

    if(!is.null(facet)){

      path_plot <- path_plot +
      facet_wrap(~ get(facet),
                 scales = "free") 
    }

    return(path_plot)

  }

  dir_wis_plot <- wis_plot(decomp_data, "directional", xmax)
  abs_wis_plot <- wis_plot(decomp_data, "absolute", xmax)

  crpsplot_rw <- horizon_path_plot(crps_scores_rw |>.d(, error_method := "directional"), "score", "CRPS by Sample", title = "CRPS - expanding window") +
    scale_y_continuous(limits = c(0.07,1.05))

ovr_assessment_plot(wis_scores, scores_cvgshort, point_scores, all_scores, "pcpi_pch")

```
\newpage

@@

```{r gdpovr, echo = FALSE, warning = TRUE, fig.dim=c(12,12)}

ovr_assessment_plot(wis_scores, scores_cvgshort, point_scores, all_scores, "ngdp_rpch")
```

>>>>

ovr_assessment_plot <- function(wis_scoredat, cvg_scoredat, point_scores, crps_scores, trgt, window_method){

  cvgscores <- cvg_scoredat |>
    .d(method == window_method) |>
    .d(target == trgt) |>
    .d(, target := NULL) |>
    .d(, method := NULL)

  decomp_data <- wis_scoredat |>
    .d(method == window_method) |>
    .d(, method := NULL)

  xmax <- max(decomp_data |> copy() |> .d(target == trgt) |> .d(, .(interval_score))) + 0.01

@@

  crps_scores_rw <- crps_scores |>
    .d(method == window_method) |>
    .d(target == trgt) |>
    .d(, target := NULL)

@@

  dir_wis_plot <- wis_plot_new(decomp_data |> .d(error_method  == "directional"), trgt, xmax = xmax, plot_name = "WIS - Directional Errors")
  abs_wis_plot <- wis_plot_new(decomp_data |> .d(error_method  == "absolute"), trgt, xmax = xmax, plot_name = "WIS - Absolute Errors")

@@

  crpsplot_rw <- horizon_path_plot(crps_scores_rw |>.d(, error_method := "directional"), "score", "CRPS by Sample", title = paste0("CRPS - ", window_method)) +
    scale_y_continuous(limits = c(0.07,1.05))

@@

ovr_assessment_plot(wis_scores, scores_cvgshort, point_scores, all_scores, trgt = "pcpi_pch", window_method = "rolling window")

```
\newpage
Some notes:
\begin{itemize}
\item Inflation: directional vs. absolute errors:
\begin{itemize}
\item difference small for IMF method, absolute slightly better, likely due to longer central intervals
\item AR and BVAR profit more from directional correction (upward bias)
\item for expanding window method, difference in coverage is smaller (-> structural breaks)

@@

```{r gdpovr_rw, echo = FALSE, warning = TRUE, fig.dim=c(12,12), warning = FALSE}

ovr_assessment_plot(wis_scores, scores_cvgshort, point_scores, all_scores, "ngdp_rpch", window_method = "rolling window")
```

@@

## Inflation
```{r infklovr_ew, echo = FALSE, warning = TRUE, fig.dim=c(12,12), warning = FALSE}

ovr_assessment_plot(wis_scores, scores_cvgshort, point_scores, all_scores, "ngdp_rpch", window_method = "expanding window")
```

@@

```{r gdpovr, echo = FALSE, warning = TRUE, fig.dim=c(12,12), warning = FALSE}

ovr_assessment_plot(wis_scores, scores_cvgshort, point_scores, all_scores, "ngdp_rpch", window_method = "expanding window")
```
